\chapter{Conclusion}\label{conclusionch}

To conclude this dissertation, we will consider the aims of both the theoretical and applied components of this project. In \textit{Chapter \ref{intro}: Introduction}, the following objectives were proposed:

\begin{itemize}
    \item Introduce the concept of this project.
    \item Provide the reader with a rounded understanding of cryptocurrencies.
    \item Explain to the reader how volatile cryptocurrency prices can be.
    \item Describe in detail the applied aspect of this project.
    \item Create a simple web application which is easy to use and clear to understand.
    \item Deliver cryptocurrency prices to the user.
    \item Provide an educated guess as to future changes in prices.
    \item Work closely with the given learning outcomes for this project.
    \item Conduct work as a team, in a professional manner akin to what is expected in industry.
\end{itemize}

Based on the findings discussed in \textit{\ref{seceval}: System Evaluation}, we can conclude that we have provided the reader with an extensive yet clear look into the world of cryptocurrency, including the defining characteristics of such an asset, driving technologies, and the innovations this field has brought. With a focus on Bitcoin, we outlined its beginnings and its rise to popularity, examining the groundbreaking blockchain technology it brought to fruition. Highlighting the viability of this new financial asset, we considered the advantages cryptocurrency brings over traditional currencies, discussing the anonymity and security in trading such an asset. However, such an asset does not come without its flaws and uncertainties, as discussed in \textit{Chapter \ref{predictingpricesch}: Predicting the Prices of Cryptocurrencies}. The influencing factors in prices and rising levels in popularity are explained, including how they have led to an extremely volatile market, with prices fluctuating in the order of thousands on a weekly, sometimes daily, basis. 

Concluding the theoretical aspects of this project, the Currency Analyser web application was introduced. A simple web application was developed to display the prices of popular cryptocurrency Bitcoin on a graph, updating every thirty seconds with new prices. Machine learning was incorporated into the project in order to provide the user with estimates as to what the closing price of Bitcoin would be for the following day. The finished web application was successfully deployed on Heroku, and can now be used by any member of the public. 

Throughout the process as a whole there have been many new experiences, owing particularly to working as a team on such an important project. There were of course disagreements along the way, all fortunately being resolved in a professional and calm manner. The experience gained throughout the development process both in terms of applied knowledge and interpersonal skills is invaluable, and will of course be reflected upon for future projects, academic or career-based. 

With the above in mind, we can conclude that the initial objectives for this project were met in some capacity or another. Both the extensive research outlined in this dissertation and observation of the Currency Analyser web application has emphasised that one cannot predict cryptocurrency prices. One can however make an educated guess as to what future prices might be. While our machine learning model is not intended to provide very accurate result, it is close to accurate as long as there is no sudden major shift in the market. Of course, this is the volatility of cryptocurrency yet again reiterating the impossibility of predicting prices.

On a final positive note, the completion of this project has left the authors with a wealth of new knowledge and experience, including the capability to critically evaluate their own work. More often than not, the development of this project and dissertation was enjoyable, inspiring the authors to aim for the highest of standards. All aspects of this final year project has challenged the team in terms of research, development, and resolving interpersonal disagreements, all of which has added to the feeling of satisfaction upon final submission of this project.

\chapter{Appendices}

\textbf{Source Code on Github: }\\ \noindent\textcolor{NavyBlue}{\url{https://github.com/rebeccabernie/CurrencyAnalyser}}\\

\noindent\textbf{Heroku Web Application Link: }\\\ \noindent\textcolor{NavyBlue}{\url{https://currencyanalyser.herokuapp.com/}}\\

\noindent\textbf{Screencast: }\\ \noindent\textcolor{NavyBlue}{\url{https://github.com/rebeccabernie/CurrencyAnalyser/blob/master/Screencast.mp4}}

